High-Performance Data Analysis on Janus using Apache Spark

### 2013-12-17 paper.tex
wrote evaluation overview, grep test overview, and runtime and speedup results

\section{Evaluation}
% janus
    % overview
    % describe what spark test where
%%% DIRECTLY COPY-PASTED TEXT %%%
University of the University of Colorado's Top-
31 (June 2010) supercomputer system, Janus
    % specs (ref)
%%% DIRECTLY COPY-PASTED TEXT %%%
The final Janus computational
system consists of 1,368 nodes, each containing two six-
core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total
of 2,736 processors and 16,416 cores with a theoretical
peak performance of 184 TFLOP/s. Each core has 2
GB of 1333 MHz DDR3 RAM for a total of 24 GB per
node and 32 TB of memory in the entire system. (REF)
% spark
    % reason for spark (diskless)
    % installation (default???)
    % version number (Spark 0.8.0, Scala 2.9.3)
% purpose of trials (speedup, efficiency, install on Janus)
    % find ideal configurations

\subsection{Testing}
% grep
    % overview
    % program source
    % grep complexity
    % file content
    % files sizes
    % node core counts
    % serial runtime
    % workers per node
% page rank
    % overview
    % ...

\subsection{Results}
% cores grep test
    % scale up, workers per node (graph)
    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeTimes.png}
        \caption{Grep runtimes with different workers per node over 20GB file}
        \label{fig:workNodeTime}
    \end{figure}
    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeSpeedup.png}
        \caption{Grep speedup with different workers per node over 20GB file}
    \end{figure}
    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeEfficiency.png}
        \caption{Grep efficiency with different workers per node over 20GB file}
        \label{fig:workNodeEff}
    \end{figure}
    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeKarpFlatt.png}
        \caption{Grep Karp-Flatt metric with different workers per node over
                20GB file}
        \label{fig:workNodeKF}
    \end{figure}
        % stays relative constant, serial is the issue
% large grep test
    % serial run time (graph)
    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/bigDataTimes.png}
        \caption{Average grep runtimes over 100GB file}
        \label{fig:bigTime}
        % stays relative constant, serial is the issue
    \end{figure}
% pagerank test 1
% pagerank test 2
% ...

\section{Discussion}
% small grep test
    % ideal balance between speedup and efficiency
% cores grep test
    % ideal number of workers per core
    % close enough not to matter
    % don't use 6-12
% large grep test
    % ideal workers per problem size
    % follow up test with delay
    % take away on IO
% pagerank test 1
% pagerank test 2
% ...

% workings and master comm of spark protocol

>>>>

\section{Evaluation}
We conducted our evaluation of the Spark framework on the University of
Colorado's Janus supercomputer.  Janus consists of 1,368 nodes, each containing
two six core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total of 12 cores
per node.  Additionally, each core has 2 GB of 1333 MHz DDR RAM, for a total of
24 GB per node \cite{JanusPaper}. In our research, we employed Spark version
0.8.0 along with Scala version 2.9.3. During our testing of the Spark framework,
we access input files from Janus' **lustre distrubted file system**. % WHAT ELSE
Our aim was to assess the Spark on Janus and determine optimal configurations
to achieve the best performance.

\subsection{Testing}
Following the example from Google's first paper on MapReduce, the first test we
performed was a parallel grep. Here an application would scan through a large
test file, looking for lines that matched a given regular expression, finally
returning the total number of matched lines. Two series of test were run on two
different files sizes. We first wanted to see how Spark scales in an ideal
setting, where the entire data set fits in memory. ***This way we avoid the
large performance hit Spark takes writing intermediate results to disk***.
Given that each node has a total of 24 GB of RAM, we limited our first problem
size to 20GB. This allowed a single node to contain the entire problem in
memory, leaving some space for the OS and other supporting processes. We
excuted this first test with 1 to 5 nodes. The second test we performed was
run with 5 to 50 nodes, each processing a 100GB file.

Ten trials were performed for each test, allows us to compute average
runtimes. To compute the speedup, Spark runtimes were compared to the UNIX grep
commands serial runtime on the same file. Finally, during our first test on the
20 GB file, we evaluated performance using a different number of workers per
node.  In Spark, a worker is an executor with it's own JVM which processes it's
own portion of the distributed data set. By default, a single worker is
employed on each node. For this test, we wanted to see if this was the ideal
setting on Janus.

% page rank
    % overview
    % ...

\subsection{Results}
In the first Grep test, we examined how Spark scales over a 20GB file using 1
to 5 nodes. Additionally, each trial was executed with a different number
workers per node. Figure \ref{fig:workNodeTime} shows the average runtimes as
the number of nodes increases. Each line in this graph represents a different
number of workers per node. We clearly can see from this graph that 6 and 12
workers per node yields significantly worse performance, while the results
produced by 1 to 3 workers is relatively equal.

    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeTimes.png}
        \caption{Grep runtimes with different workers per node over 20GB file}
        \label{fig:workNodeTime}
    \end{figure}

Figure \ref{fig:workNodeSpeed} illustrates the results in terms of speedup.
Here we are comparing the resulting runtimes to the UNIX grep serial runtime of
1,120 seconds, processing the same 20 GB file. Note that here 1 node represents
12 cores.  So the initial speedup between 3 and 4 is not as impressive as one
might expect.  We also see that the speedup begins to level out around 4 nodes,
and in the case of 1 to 3 workers per node, it even decreases from 4 to 5
nodes.

    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeSpeedup.png}
        \caption{Grep speedup with different workers per node over 20GB file}
        \label{fig:workNodeSpeed}
    \end{figure}

% in graphRef we can see how well the cores are being utilized
% in an efficiency graph
% as expected from previous graphs, efficiency not great
% starts under 40% efficiency with 1 node
% end up at just about 10% efficiency with 5 nodes, or 60 cores

    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeEfficiency.png}
        \caption{Grep efficiency with different workers per node over 20GB file}
        \label{fig:workNodeEff}
    \end{figure}

% finally we see this data in terms of karp-flatt metric in graphRef
% with the exception of the 12 workers per node in pink
% we see that the karp-flatt metric remains fairly constaint
% values between 0.12 and 0.14
% this constant value would suggest that the serial runtime
% is what is preventing better scaling

    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/workerPerNodeKarpFlatt.png}
        \caption{Grep Karp-Flatt with different workers per node over 20GB file}
        \label{fig:workNodeKF}
    \end{figure}

% to see if we could see better scaling over a largest file
% switch from 20GB to 100GB file
% accessed file from lustre drive
% as 100GB cannot fit in memory of one node, start with 5
% graphRef should the runtimes over this large file
% clearly there appears to be no change in runtimes
% remains round 500 seconds between 5 nodes and 50
% suspecting short compute time of grep expression
% switch to a more computational expensive and iterative PageRank

    \begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/bigDataTimes.png}
        \caption{Average grep runtimes over 100GB file}
        \label{fig:bigTime}
    \end{figure}

% pagerank test 1
% pagerank test 2
% ...

\section{Discussion}
% the results for the grep test over the 20gb file
% gave us some insight on how well such a problem scale in spark
% for that given task and problem size
% 3 nodes appears to be a good balance between speedup and efficiency
% we all saw how workers per node effects runtime
% while 1-3 workers remains close, clear 6-12 is not ideal
% it seems as if one my observer slightly better results
% with 2-3 nodes over 1 node, but not by much

% in the large grep test over the 100 gb file
% suprisingly obvserved no change in performance
% perhaps the results of karp-flatt metric in previous test
% gives us same insight as to why we are not seeing improvement
% we suspect the long load times are are bigger
% than the relatively short time needed to graph the loaded information
% the long load times may suggest the file is not read in parallel

% pagerank test 1
% pagerank test 2
% ...

% other problems exhibited
    % easy to install, hard to configure/profile/understand results
    % sporadic runtimes

% workers and master communication over spark protocol
